# ---
# output:
#   bookdown::pdf_document2:
#     includes:
#       in_header: latex/preamble.tex
#       before_body: latex/titlepage.tex
#     pandoc_args:
#     - --csl
#     - references/apa.csl
#   bookdown::html_document2:
#     pandoc_args:
#     - --csl
#     - references/apa.csl
#   bookdown::word_document2:
#     pandoc_args:
#     - --csl
#     - references/apa.csl
# toc-title: Table of Contents
# bibliography: references/references.bib
# link-citations: yes
# header-includes:
# - \usepackage{fancyhdr}
# - \pagestyle{fancy}
# - \renewcommand{\headrulewidth}{0pt}
# - \fancyfoot[C]{}
# - \fancyfoot[R]{\thepage}
# ---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE)
packages <- c("bookdown", "dplyr", "ggplot2", "papeR", "kableExtra", "mgcv")
package.check <- lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
        install.packages(x, dependencies = TRUE)
        library(x, character.only = TRUE)
    }
})
```
\fancyhead[LR]{}
\pagenumbering{roman}

\newpage
\cleardoublepage
\pagenumbering{arabic}
\fancyhead[L]{Employee Attrition Model}
\fancyhead[R]{Machine Learning I}
# Introduction

Employees, according to @IBM_Developer_Challenge_2020, are the foundation of any business. Its success is largely determined by the quality of its employees and their ability to stay with the company. Organizations confront a number of issues as a result of staff attrition:

1. Training new personnel is costly in terms of both money and time.
2. Potential to lose experienced employees
3. Productivity impact
4. Profitability impact

Therefore, IBM data scientists created a fictitious data set as a challenge for data scientists. Among the data types are metrics such as education level, job satisfaction, and commute distance. The dataset can be found on the company's GitHub account [@IBM_Data_2019].

# Methodology

The following topics are layed out through out this paper:

1. Linear Models
2. Extending the Linear Model: Non-linearity
3. Extending the Linear Model: Generalised Linear Models
4. Support Vector Machines
5. Neural Networks
6. Optimisation

\newpage
# Data preparation
```{r import-data, echo = FALSE, results = FALSE}
# Working directory
setwd("~/HSLU/FS_22/Machine_learning/Groupwork/Github")
emp_attrition <- read.csv("emp_attrition.csv")
head(emp_attrition)
```

## Data Transformation and Sanity Check

The code for this part is left out from the PDF due to its length...
```{r formatting-data, echo = FALSE, results = FALSE}
# Age
names(emp_attrition)[names(emp_attrition) == "ï..Age"] <- "Age"
summary(emp_attrition$Age)

# Attrition
emp_attrition$Attrition <- as.factor(emp_attrition$Attrition)
unique(emp_attrition$Attrition)

# BusinessTravel
emp_attrition$BusinessTravel <- as.factor(emp_attrition$BusinessTravel)
levels(emp_attrition$BusinessTravel) <- c("Rarely", "Frequently", "None")
emp_attrition$BusinessTravel <- relevel(emp_attrition$BusinessTravel, ref = "None")
unique(emp_attrition$BusinessTravel)

# DailyRate
summary(emp_attrition$DailyRate)

# Department
emp_attrition$Department <- as.factor(emp_attrition$Department)
levels(emp_attrition$Department) <- c("Sales", "R&D", "HR")
unique(emp_attrition$Department)

# DistanceFromHome
summary(emp_attrition$DistanceFromHome)

# Education
emp_attrition$Education <- as.factor(emp_attrition$Education)
levels(emp_attrition$Education) <- c("Below College", "College", "Bachelor", "Master", "Doctor")
unique(emp_attrition$Education)

# EducationField
emp_attrition$EducationField <- as.factor(emp_attrition$EducationField)
unique(emp_attrition$EducationField)

# EmployeeCount
unique(emp_attrition$EmployeeCount) # represents headcount which for all employee's in this dataset 1

# EmployeeNumber
summary(emp_attrition$EmployeeNumber) # internal employee id

# EnvironmentSatisfaction
emp_attrition$EnvironmentSatisfaction <- as.factor(emp_attrition$EnvironmentSatisfaction)
levels(emp_attrition$EnvironmentSatisfaction) <- c("Low", "Medium", "High", "Very High")
unique(emp_attrition$EnvironmentSatisfaction)

# Gender
emp_attrition$Gender <- as.factor(emp_attrition$Gender)
unique(emp_attrition$Gender)

# HourlyRate
summary(emp_attrition$HourlyRate)

# JobInvolvement
emp_attrition$JobInvolvement <- as.factor(emp_attrition$JobInvolvement)
levels(emp_attrition$JobInvolvement) <- c("Low", "Medium", "High", "Very High")
unique(emp_attrition$JobInvolvement)

# JobLevel
emp_attrition$JobLevel <- as.factor(emp_attrition$JobLevel)
unique(emp_attrition$JobLevel)

# JobRole
emp_attrition$JobRole <- as.factor(emp_attrition$JobRole)
unique(emp_attrition$JobRole)

# JobSatisfaction
emp_attrition$JobSatisfaction <- as.factor(emp_attrition$JobSatisfaction)
levels(emp_attrition$JobSatisfaction) <- c("Low", "Medium", "High", "Very High")
unique(emp_attrition$JobSatisfaction)

# MaritalStatus
emp_attrition$MaritalStatus <- as.factor(emp_attrition$MaritalStatus)
unique(emp_attrition$MaritalStatus)

# MonthlyIncome
summary(emp_attrition$MonthlyIncome)

# MonthlyRate
summary(emp_attrition$MonthlyRate)

# NumCompaniesWorked
unique(emp_attrition$NumCompaniesWorked)

# Over18
unique(emp_attrition$Over18) # all employee's are above 18

# OverTime
emp_attrition$OverTime <- as.factor(emp_attrition$OverTime)
unique(emp_attrition$OverTime)

# PercentSalaryHike
summary(emp_attrition$PercentSalaryHike)

# PerformanceRating
emp_attrition$PerformanceRating <- as.factor(emp_attrition$PerformanceRating)
levels(emp_attrition$PerformanceRating) <- c("Excellent", "Outstanding")
unique(emp_attrition$PerformanceRating) # only employee's who perform 3 or 4

# RelationshipSatisfaction
emp_attrition$RelationshipSatisfaction <- as.factor(emp_attrition$RelationshipSatisfaction)
levels(emp_attrition$RelationshipSatisfaction) <- c("Low", "Medium", "High", "Very High")
unique(emp_attrition$RelationshipSatisfaction)

# StandardHours
unique(emp_attrition$StandardHours) # the StandardHours for all employee's is 80, after check this means that this data has a 9/80 work schedule. Hence, employees work 80 hours in 9 days. So not a standard as 5/42 as in switzerland..

# StockOptionLevel
emp_attrition$StockOptionLevel <- as.factor(emp_attrition$StockOptionLevel)
unique(emp_attrition$StockOptionLevel)

# TotalWorkingYears
summary(emp_attrition$TotalWorkingYears)

# TrainingTimesLastYear
unique(emp_attrition$TrainingTimesLastYear)

# WorkLifeBalance
emp_attrition$WorkLifeBalance <- as.factor(emp_attrition$WorkLifeBalance)
levels(emp_attrition$WorkLifeBalance) <- c("Bad", "Good", "Better", "Best")
unique(emp_attrition$WorkLifeBalance)

# YearsAtCompany
summary(emp_attrition$YearsAtCompany)

# YearsInCurrentRole
summary(emp_attrition$YearsInCurrentRole)

# YearsSinceLastPromotion
summary(emp_attrition$YearsSinceLastPromotion)

# YearsWithCurrManager
summary(emp_attrition$YearsWithCurrManager)

labels(emp_attrition) <- colnames(emp_attrition)
labels(emp_attrition, which = c("Age", "BusinessTravel", "DailyRate", "DistanceFromHome", "EducationField", "EmployeeNumber", "EnvironmentSatisfaction", "HourlyRate", "JobInvolvement", "JobLevel", "JobRole", "JobSatisfaction", "MaritalStatus", "MonthlyIncome", "MonthlyRate", "NumCompaniesWorked", "OverTime", "PercentSalaryHike", "PerformanceRating", "RelationshipSatisfaction", "StockOptionLevel", "TotalWorkingYears", "TrainingTimesLastYear", "WorkLifeBalance", "YearsAtCompany", "YearsInCurrentRole", "YearsSinceLastPromotion", "YearsWithCurrManager")) <- c("Age (years)", "Business Travel", "Daily Rate (salary per day)", "Distance From Home (1-29)", "Education Field", "Employee ID", "Environment Satisfaction", "Hourly Rate (salary per hour)", "Job Involvement", "Job Level (internal company hierachy 1-5)", "Job Role", "Job Satisfaction", "Marital Status", "Monthly Income", "Monthly Rate (salary per month)", "Number of Companies worked before starting this position", "has done over time", "% Salary Hike", "Performance Rating", "Relationship Satisfaction", "Stock Option Level (number of stocks an employee owns from 0-3)", "Total Working Years", "Number Of Trainings Done Last Year", "Work Life Balance", "Number Of Years At Company", "Number of Years In Current Role", "Number Of Years Since Last Promotion", "Number Of Years With Current Manager")
```

## Data Cleaning

```{r data-cleaning}
emp_attrition <- subset(emp_attrition, select = -c(EmployeeCount, StandardHours, Over18))
```

- EmployeeCount (represents the head count which is 1 for all employee, hence drop this)
- StandardHours (StandardHours for all employee's is 80, therefore this data has a 9/80 work schedule. Hence, employees work 80 hours in 9 days. So not a standard as 5/42 as in switzerland, we drop this)
- Over18 (all employee's are 18 or above and it's capured in age, hence drop this variable)

## Missing Value Check
```{r check-for-missing-values, echo = TRUE, results = FALSE}
# Do we have any missing values?
sapply(emp_attrition, function(x) all(is.na(x) | x == '' ))
```
There are no missing values in this dataset.

\newpage
## Overview of Dataset

```{r Overview, echo = TRUE, results = TRUE}
str(emp_attrition)
```
Interesting variables:

- Attrition (yes/no)
- Departement (3 levels)
- DistanceFromHome(continous)
- JobInvolvement (4 levels)
- JobSatisfaction (4 levels)
- Monthly income (continous)
- NumCompaniesWorked (continous)
- TrainingTimesLastYear (continous)
- WorkLifeBalance (4 levels)
- YearsAtCompany (continous) #might be used for LM model
- YearWithCurrManager

\newpage
# Exploration


\newpage

# Linear Regression
```{r Linear Regression}
# All variables without log transformation
lm.YearsAtCompany.0 <- lm(formula = YearsAtCompany ~ ., data = emp_attrition)
lm.summary.0 <- summary(lm.YearsAtCompany.0)
lm.summary.0

#Histogram plotsloop
emp_attrition_numerical_only <- emp_attrition[, c(1, 4, 6, 12, 18, 19, 20, 26,
                                                  27, 29, 30, 31, 32)]
names_numerical <- colnames(emp_attrition_numerical_only)

for (i in 1:ncol(emp_attrition_numerical_only)) {
  variable <- emp_attrition_numerical_only[, i]
  hist(variable, breaks = 30, col = "blue", main = names_numerical[i])
}

# Log transformation of variables that should only have positive values (or 0)
emp_attrition_log <- emp_attrition
#emp_attrition_log["DailyRate"] <- log(emp_attrition_log["DailyRate"]) 
emp_attrition_log["DistanceFromHome"] <- log(emp_attrition_log["DistanceFromHome"]) 
#emp_attrition_log["HourlyRate"] <- log(emp_attrition_log["HourlyRate"])
emp_attrition_log["MonthlyIncome"] <- log(emp_attrition_log["MonthlyIncome"])
emp_attrition_log["YearsAtCompany"] <- log(emp_attrition_log["YearsAtCompany"])
emp_attrition_log["NumCompaniesWorked"] <- log(emp_attrition_log["NumCompaniesWorked"])
emp_attrition_log["TotalWorkingYears"] <- log(emp_attrition_log["TotalWorkingYears"])
#emp_attrition_log["TrainingTimesLastYear"] <- log(emp_attrition_log["TrainingTimesLastYear"])
emp_attrition_log["YearsInCurrentRole"] <- log(emp_attrition_log["YearsInCurrentRole"])
emp_attrition_log["YearsSinceLastPromotion"] <- log(emp_attrition_log["YearsSinceLastPromotion"])
emp_attrition_log["YearsWithCurrManager"] <- log(emp_attrition_log["YearsWithCurrManager"])


# Na's / -Inf values check after log transformation # Back to 0
sum(is.na(emp_attrition_log)) # 0
sum(emp_attrition_log == "-Inf") #1340
emp_attrition_log[emp_attrition_log == "-Inf"] <- 0
emp_attrition_log[emp_attrition_log == "NaN"] <- 0


# All variables with log transformation
lm.YearsAtCompany.1 <- lm(formula = YearsAtCompany ~ ., data = emp_attrition_log)
lm.summary.1 <- summary(lm.YearsAtCompany.1)
lm.summary.1
```
Findings LM Models _log:
- Variable Age, p-value slightly > 0.05 - but is kept
- Differences between Departements / Roles / Education etc. might be analysed in detail (different significance levels)
- For next model only general variables with p-value > 0.05 are kept, excluding 
variables with factor levels where not all levels are significant. 
- Adjusted R-squared:  0.7861 without log transformation / with log transformation 0.85  

Findings Histograms:
- Variables that show right skeewedness are log transformed. 

# Residual analysis LM Models

```{r}
# Without log transformation
# get list of residuals
lm.YearsAtCompany.0_residuals <- resid(lm.YearsAtCompany.0)

#produce residual vs. fitted plot
plot(fitted(lm.YearsAtCompany.0), lm.YearsAtCompany.0_residuals)
abline(0,0) #add a horizontal line at 0 

#create Q-Q plot for residuals
qqnorm(lm.YearsAtCompany.0_residuals)
qqline(lm.YearsAtCompany.0_residuals)#add a straight diagonal line to the plot
 
#Create density plot of residuals
plot(density(lm.YearsAtCompany.0_residuals))


# With log transformation
# get list of residuals
lm.YearsAtCompany.1_residuals <- resid(lm.YearsAtCompany.1)

#produce residual vs. fitted plot
plot(fitted(lm.YearsAtCompany.1), lm.YearsAtCompany.1_residuals)
abline(0,0) #add a horizontal line at 0 

#create Q-Q plot for residuals
qqnorm(lm.YearsAtCompany.1_residuals)
qqline(lm.YearsAtCompany.1_residuals)#add a straight diagonal line to the plot
 
#Create density plot of residuals
plot(density(lm.YearsAtCompany.1_residuals))
```
Without log transformation:
-  Resiudals vs. fitted plot: The spread of residuals seems to be higher for higher fitted values. 
Therefore, the normal distributions should be further verified. 
- QQ-Plot: The strayed residuals around the tails indicate that they might not 
be normally distributed.
- Density plot: As the residuals are mainly bell shaped, only displaying long
tails, the normal distribution will still be assumed. Nevertheless, additional
models will be fitted going forward.

With log transformation:
- Resiudals vs. fitted plot: The spread of residuals was reduced compared to the 
model without log transformation.
- QQ-Plot: In this plot the residuals seem better distributed but still strayed
around higher values.
- Density plot: The bell shape is still given, with less long tails. 


```{r Linear Regression selection}
# Significant variables, excluding variables where not all factor levels are 
# significant for simplicity
lm.YearsAtCompany.2 <- lm(YearsAtCompany ~ 
                            Age +
                            NumCompaniesWorked +
                            PercentSalaryHike +
                            TotalWorkingYears +
                            YearsInCurrentRole +
                            YearsSinceLastPromotion +
                            YearsWithCurrManager
                            , data = emp_attrition_log)
lm.summary.2 <- summary(lm.YearsAtCompany.2)
lm.summary.2
```
Findings:
- All variables display a p-value > 0.05. Adjusted R-squared:  0.8467 - only a 
0.0033 difference to the previous model including all the variables. 
Interpretation:
- Intercept: YearsAtCompany if all other variables were 0. Given variables such 
as Age, it is pointless to interpret the intercept in this model.  
- Age: With an increase of 1 year of age, the dependent variable YearsAtCompany 
decreases by 0.003173. Seems logical, if people e.g. are going to retire 1 year earlier?
- NumCompaniesWorked: If the variable increases by 1, the YearsAtCompany decrease 
by 0.159989. People who change employers a lot might stick with this pattern.
- PercentSalaryHike: If the Salary is increased by 1%, the YearsAtCompany decrease
by 0.007561. This seems controversial but is also a sign that the salary-system 
maybe does not work ideal. 
- TotalWorkingYears: If the TotalWorkingYears increase by 1, the YearsAtCompany
increase by 0.312428.
- YearsInCurrentRole: If the YearsInCurrentRole increase by 1, the YearsAtCompany
increase by 0.453442.
- YearsSinceLastPromotion: If the YearsSinceLastPromotion increase by 1,
the YearsAtCompany increase by 0.044885. This might suggest that people who expect
a promotion are willing to keep working at the company.
- YearsWithCurrManager: If the YearsWithCurrManager increase by 1, the YearsAtCompany
increase by 0.385965. This hints that employees like consitency in the managment.

```{r}
# Including variable with at least one significant factor level # Exclude?
lm.YearsAtCompany.3 <- lm(YearsAtCompany ~ 
                            Age +
                            Education +
                            EnvironmentSatisfaction + 
                            JobRole +
                            JobSatisfaction +
                            NumCompaniesWorked +
                            PercentSalaryHike +
                            TotalWorkingYears +
                            YearsInCurrentRole +
                            YearsSinceLastPromotion +
                            YearsWithCurrManager
                            , data = emp_attrition_log)
lm.summary.3 <- summary(lm.YearsAtCompany.3)
lm.summary.3
```

## Generalised Linear Models

\newpage
# Conclusion

\newpage
# Session Information {-}

```{r session-info}
sessionInfo()
```

\newpage
# References

