---
output:
  bookdown::pdf_document2:
    includes:
      in_header: latex/preamble.tex
      before_body: latex/titlepage.tex
    pandoc_args:
    - --csl
    - references/apa.csl
  bookdown::html_document2:
    pandoc_args:
    - --csl
    - references/apa.csl
  bookdown::word_document2:
    pandoc_args:
    - --csl
    - references/apa.csl
toc-title: Table of Contents
bibliography: references/references.bib
link-citations: yes
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \renewcommand{\headrulewidth}{0pt}
- \fancyfoot[C]{}
- \fancyfoot[R]{\thepage}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE)
packages <- c("bookdown", "dplyr", "ggplot2")
package.check <- lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
        install.packages(x, dependencies = TRUE)
        library(x, character.only = TRUE)
    }
})
```
\fancyhead[LR]{}
\pagenumbering{roman}

\newpage
\cleardoublepage
\pagenumbering{arabic}
\fancyhead[L]{Example Model}
\fancyhead[R]{Machine Learning I}
# Introduction

Employees, according to @IBM_Developer_Challenge_2020, are the foundation of any business. Its success is largely determined by the quality of its employees and their ability to stay with the company. Organizations confront a number of issues as a result of staff attrition:

1. Training new personnel is costly in terms of both money and time.
2. Potential to lose experienced employees
3. Productivity impact
4. Profitability impact

Therefore, IBM data scientists created a fictitious data set as a challenge for data scientists. Among the data types are metrics such as education level, job satisfaction, and commute distance. The dataset can be found on the company's GitHub account [@IBM_Data_2019].

```{r}
emp_attrition <- read.csv("./data/emp_attrition.csv")
head(emp_attrition)
```

# Methodology

# Data preparation

# Conclusion

\newpage
# References
